---
title: "Getting Started with bayescoveragemodel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with bayescoveragemodel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

This vignette demonstrates a quick example workflow for fitting Bayesian
transition models to health coverage indicators. We use minimal sampling
iterations for demonstration purposes.

For full model fitting with proper sampling settings, see the
[Global Model Fitting](global_model_fitting.html) article.

## Setup

```{r setup, message=FALSE, warning=FALSE}
library(bayescoveragemodel)
library(dplyr)
library(ggplot2)
```

## Load Data

The package includes example survey data for health coverage indicators
(ANC4, institutional delivery, vaccination).

```{r load-data}
# Load survey data
data_folder <- system.file("extdata", package = "bayescoveragemodel")

# If using development version with data_raw folder
if (data_folder == "") {
  data_folder <- "data_raw"
}

# Load regions metadata
regions_dat <- readr::read_csv(
  file.path(data_folder, "regions_updated.csv"),
  show_col_types = FALSE
)
```

## Process Data

Use `process_data()` to prepare the survey data for model fitting.

```{r process-data, eval=FALSE}
# Load raw survey data (example using haven for .dta files)
dat0 <- haven::read_dta(file.path(data_folder, "ICEH_national.dta"))

# Process data for a specific indicator
dat <- process_data(
  dat = dat0,
  regions_dat = regions_dat,
  indicator = "anc4"
)

head(dat)
```

## Explore Data

Before fitting models, explore the survey data using `explore_data()`.

```{r explore-data, eval=FALSE}
# Create exploration plots
plots <- explore_data(

data = dat,
  indicator_col = "indicator",
  group_col = "country",
  data_types_col = "data_series_type",
  indicator_name = "ANC4",
  region_col = "cluster"
)

# View one of the plots
plots[[1]]
```

## Fit Model (Test Settings)

Fit a model using test settings with minimal iterations. This runs quickly
but should not be used for inference.

```{r fit-model, eval=FALSE}
# Write Stan models (run once)
write_model()

# Fit model with test settings (fast, not for inference)
fit <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  runnumber = "test_run",
  chains = 2,
  # Test settings for quick demonstration
  iter_sampling = 10,
  iter_warmup = 5,
  get_posteriors = TRUE
)
```

## Plot Results

Visualize the fitted estimates.

```{r plot-results, eval=FALSE}
# Plot estimates for all countries
p <- plot_estimates_local_all(
  results = fit,
  save_plots = FALSE,
  indicator_name = "anc4"
)

# Display a plot
p[[1]]
```

## Next Steps

For production model fitting with proper sampling:

1. **Global Model Fitting**: See the [Global Model Fitting](global_model_fitting.html)
   article for the complete workflow with recommended sampling settings.

2. **Validation**: See the [Out-of-Sample Validation](validation.html) article
   for assessing predictive performance.

3. **Diagnostics**: See the [Model Diagnostics](diagnostics.html) article for
   convergence checks and hierarchical model assessment.
