---
title: "Model Diagnostics"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article describes how to assess model fit and convergence for
Bayesian transition models.

## Setup

```{r setup}
library(bayescoveragemodel)
library(dplyr)
library(ggplot2)
library(localhierarchy)
```

## Load Fitted Models

Load previously fitted models for diagnostics:

```{r load-fits}
# Choose indicator
indicator <- "anc4"

# Load step 1a and 1b fits
fit1a <- get_fit(indicator, runstep = "step1a", folder_suffix = "run_workflow")
fit1b <- get_fit(indicator, runstep = "step1b", folder_suffix = "run_workflow")
```

## Visualize Estimates

Create plots for all countries to visually assess model fit:

```{r plot-estimates}
# Plot estimates from step 1a
plots_fit1a <- plot_estimates_local_all(
  results = fit1a,
  save_plots = TRUE,
  indicator_name = indicator
)

# Plot estimates from step 1b
plots_fit1b <- plot_estimates_local_all(
  results = fit1b,
  save_plots = TRUE,
  indicator_name = indicator
)
```

## Convergence Diagnostics

Check MCMC convergence using standard diagnostics:

```{r convergence}
# Get convergence diagnostics for step 1a
diag1a <- get_convergence_diagnostics(fit1a)
print(diag1a)

# Get convergence diagnostics for step 1b
diag1b <- get_convergence_diagnostics(fit1b)
print(diag1b)
```

Key diagnostics to check:

-   **Rhat**: Should be close to 1.0 (\< 1.01 is good)

-   **ESS (Effective Sample Size)**: Should be sufficiently large

-   **Divergent transitions**: Should be 0 or very few

## Hierarchical Model Checks

Assess the hierarchical structure of the model:

```{r hier-checks}
# Plot hierarchical checks for step 1a
plot_hierchecks(fit1a)

# Plot hierarchical checks for step 1b
plot_hierchecks(fit1b)
```

These plots help assess:

-   Prior-posterior comparison for hierarchical hyperparameters

-   Shrinkage behavior across groups

-   Model assumptions about parameter distributions

## Compare Model Fits

Compare estimates from different model runs:

```{r compare-fits}
# Compare step 1a and 1b fits
comparison <- compare_fits(fit1a, fit1b)
```

## Diagnostics for Multiple Indicators

Run diagnostics across all indicators:

```{r all-indicators}
indicators <- c("anc4", "ideliv", "vdpt")

for (ind in indicators) {
  cat("\n=== Diagnostics for", ind, "===\n")

  # Load fit
  fit <- get_fit(ind, runstep = "step1b", folder_suffix = "run_workflow")

  # Check convergence
  diag <- get_convergence_diagnostics(fit)
  print(diag)

  # Create plots
  plot_estimates_local_all(fit, save_plots = TRUE, indicator_name = ind)
  plot_hierchecks(fit)
}
```

## Next Steps

-   [Out-of-Sample Validation](validation.html): Assess predictive
    performance
