---
title: "Model Diagnostics"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article describes how to assess model fit and convergence for
Bayesian transition models.

## Setup

```{r setup}
library(bayescoveragemodel)
library(dplyr)
library(ggplot2)
library(localhierarchy)
```

## Load Fitted Models

Load previously fitted models for diagnostics:

```{r load-fits}
# Choose indicator and model type
indicator <- "anc4"
model_name <- "spline"  # or "rw2"

# Load step 1ab fit (main global model)
fit <- get_fit(indicator, runstep = "step1ab", folder_suffix = "run_workflow")

# Load validation fit
fit_val <- get_fit(indicator, runstep = "step1ab", folder_suffix = "val_workflow")
```

## Visualize Estimates

Create plots for all countries to visually assess model fit:

```{r plot-estimates}
# Plot estimates from main fit
plots_main <- plot_estimates_local_all(
  results = fit,
  save_plots = TRUE,
  indicator_name = indicator
)

# Compare main fit vs validation
plots_comparison <- plot_estimates_local_all(
  results = fit,
  results2 = fit_val,
  modelnames = c("Main", "Validation"),
  save_plots = TRUE,
  indicator_name = indicator
)
```

## Convergence Diagnostics

Check MCMC convergence using standard diagnostics:

```{r convergence}
# Get convergence diagnostics
diag <- get_convergence_diagnostics(fit)
print(diag)
```

Key diagnostics to check:

-   **Rhat**: Should be close to 1.0 (\< 1.01 is good)

-   **ESS (Effective Sample Size)**: Should be sufficiently large

-   **Divergent transitions**: Should be 0 or very few

## Hierarchical Model Checks

Assess the hierarchical structure of the model:

```{r hier-checks}
# Plot hierarchical checks
plot_hierchecks(fit)
```

These plots help assess:

-   Prior-posterior comparison for hierarchical hyperparameters

-   Shrinkage behavior across groups

-   Model assumptions about parameter distributions

## Compare Model Fits

Compare estimates from different model runs (e.g., spline vs RW2):

```{r compare-fits}
# Load both model types
fit_spline <- get_fit(indicator, runstep = "step1ab", folder_suffix = "run_spline")
fit_rw2 <- get_fit(indicator, runstep = "step1ab", folder_suffix = "run_rw2")

# Compare spline and RW2 models
plots_model_comparison <- plot_estimates_local_all(
  results = fit_spline,
  results2 = fit_rw2,
  modelnames = c("Spline", "RW2"),
  save_plots = TRUE,
  indicator_name = indicator
)
```

## Diagnostics for Multiple Indicators

Run diagnostics across all indicators:

```{r all-indicators}
indicators <- c("anc4", "ideliv", "vdpt")

for (ind in indicators) {
  cat("\n=== Diagnostics for", ind, "===\n")

  # Load fit
  fit <- get_fit(ind, runstep = "step1ab", folder_suffix = "run_workflow")

  # Check convergence
  diag <- get_convergence_diagnostics(fit)
  print(diag)

  # Create plots
  plot_estimates_local_all(fit, save_plots = TRUE, indicator_name = ind)
  plot_hierchecks(fit)
}
```

## Next Steps

-   [Out-of-Sample Validation](validation.html): Assess predictive
    performance
