---
title: "Global Model Fitting"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article describes the complete workflow for fitting global Bayesian
transition models to health coverage indicators. Global models estimate
hierarchical parameters across all countries, which are then used for
country-specific (local) estimation.

## Setup

```{r setup}
library(bayescoveragemodel)
library(dplyr)
library(ggplot2)
library(haven)
library(cmdstanr)
library(localhierarchy)
```

## Load and Process Data

```{r load-data}
# Read national survey data and region metadata
data_folder <- "data_raw"
dat0 <- read_dta(here::here(data_folder, "ICEH_national.dta"))
regions_dat <- readr::read_csv(
  here::here(data_folder, "regions_updated.csv"))

# Choose an indicator: "anc4", "ideliv", or "vdpt"
indicator_select <- "anc4"

# Process data
dat <- process_data(
  dat = dat0,
  regions_dat = regions_dat,
  indicator = indicator_select
)
```

## Write Stan Models

Before fitting, ensure the Stan models is written:
(We can remove this step if shipped with package)
```{r write-models}
# Run once to compile Stan models
write_model()
```

## Choosing a Model

The `model_name` argument in `fit_model()` determines which temporal model is used:
- **`"spline"`** (default): Transition model, ARIMA(1,1,0) with spline-based drift term.

- **`"rw2"`**: Local rate of change, ARIMA(1,2,0).

```{r model-choice}
# Using the spline model (default)
fit_spline <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  model_name = "spline",  # default, can be omitted
  chains = 4
)

# Using the RW2 model
fit_rw2 <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  model_name = "rw2",
  runnumber = "rw2",
  chains = 4
)
```

The choice between models depends on the application. The RW2 model tends to
produce smoother estimates while the spline model offers more flexibility in
capturing non-linear trends.

## Option 1: Sequential Fitting (Steps 1a and 1b)
This approach fits the model in two stages:- **Step 1a**: Fit without smoothing terms to identify outliers- **Step 1b**: Fit with smoothing terms using 1a results

### Step 1a: Initial Fit

```{r fit-1a}
fit1a <- fit_model(
  runstep = "step1a",
  survey_df = dat,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  get_posteriors = TRUE,
  chains = 4
)
```

### Identify Outliers

```{r identify-outliers}
# Find possible outliers in step 1a
fit1a$outliers <- identify_outliers_in_global_fit(fit1a)

# Prepare fit object for step 1b
fit <- fit1a
fit$samples <- NULL

# Assign outlier flags to data
dat3 <- assign_outliers(dat, outlier_record_ids = fit$outliers)
```

### Step 1b: Fit with Smoothing

```{r fit-1b}
fit1b <- fit_model(
  runstep = "step1b",
  global_fit = fit,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  survey_df = dat3,
  get_posteriors = TRUE,
  chains = 4
)
```

## Option 2: Combined Fitting (Step 1ab)

For a simpler workflow, use step 1ab which combines both steps:
```{r fit-1ab}
fit1ab <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  get_posteriors = TRUE,
  chains = 4
)
```

## Local Fitting for All Countries

After global fitting, run local models for all countries to obtain final estimates:
```{r local-fitting}
fit_local_global <- fit_model(
  runstep = "local_national",
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  survey_df = dat,
  save_post_summ = TRUE,
  get_posteriors = TRUE,
  chains = 4
)
```

## Visualize Results

```{r plot-results}
# Plot estimates for all countries
p <- plot_estimates_local_all(
  results = fit_local_global,
  save_plots = TRUE,
  indicator_name = indicator_select
)
```

## Sampling Settings

For production runs, use default sampling settings (or increase for better convergence):
```{r sampling-settings}
# Default settings (recommended for production)
fit <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  chains = 4
  # iter_sampling = 1000 (default)
  # iter_warmup = 1000 (default)
)
```

For testing or quick exploration, use reduced settings:
```{r test-settings}
# Test settings (NOT for inference)
fit <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  chains = 2,
  iter_sampling = 10,
  iter_warmup = 5
)
```

## Output Storage

Model fits are stored in a `bayestransitionoutput` folder one level up from
the project root. Use `get_fit()` to load saved fits:
```{r load-fit}
# Load a previously saved fit
fit <- get_fit(
  indicator = "anc4",
  runstep = "step1b",
  folder_suffix = "run_workflow"
)
```

## Next Steps

- [Out-of-Sample Validation](validation.html): Assess predictive performance
- [Model Diagnostics](diagnostics.html): Check convergence and model fit
