---
title: "Global Model Fitting"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article describes the complete workflow for fitting global Bayesian
transition models to health coverage indicators. Global models estimate
hierarchical parameters across all countries, which are then used for
country-specific (local) estimation.

## Setup

```{r setup}
library(bayescoveragemodel)
library(dplyr)
library(ggplot2)
library(haven)
library(cmdstanr)
library(localhierarchy)
```

## Load and Process Data

```{r load-data}
# Read national survey data and region metadata
data_folder <- "data_raw"
dat0 <- read_dta(here::here(data_folder, "ICEH_national.dta"))
regions_dat <- readr::read_csv(
  here::here(data_folder, "regions_updated.csv"))

# Choose an indicator: "anc4", "ideliv", or "vdpt"
indicator_select <- "anc4"

# Process data
dat <- process_data(
  dat = dat0,
  regions_dat = regions_dat,
  indicator = indicator_select
)
```



## Global Model Fitting (Step 1ab)

The primary workflow uses step 1ab to fit the global model:
```{r fit-1ab}
fit1ab <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  get_posteriors = TRUE,
  chains = 4
)
```

## Step 1a for Residual Extraction (Optional)

Step 1a can be run separately to extract residuals for model checking. 

```{r fit-1a}
fit1a <- fit_model(
  runstep = "step1a",
  survey_df = dat,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  get_posteriors = TRUE,
  chains = 4
)

# Extract residuals
residuals <- get_residuals_samples(fit = fit1a)
```

## Local Fitting for All Countries

After global fitting, run local models for all countries to obtain final estimates:
```{r local-fitting}
fit_local_global <- fit_model(
  runstep = "local_national",
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  survey_df = dat,
  save_post_summ = TRUE,
  get_posteriors = TRUE,
  chains = 4
)
```

## Visualize Results

```{r plot-results}
# Plot estimates for all countries
p <- plot_estimates_local_all(
  results = fit_local_global,
  save_plots = TRUE,
  indicator_name = indicator_select
)
```



## Output Storage

Model fits are stored in a `bayestransitionoutput` folder one level up from
the project root. Use `get_fit()` to load saved fits:
```{r load-fit}
# Load a previously saved fit
fit <- get_fit(
  indicator = "anc4",
  runstep = "step1b",
  folder_suffix = "run_workflow"
)
```

## Next Steps

- [Out-of-Sample Validation](validation.html): Assess predictive performance
- [Model Diagnostics](diagnostics.html): Check convergence and model fit
