---
title: "Out-of-Sample Validation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article describes how to perform out-of-sample validation to assess the
predictive performance of Bayesian transition models.

## Setup

```{r setup}
library(bayescoveragemodel)
library(dplyr)
library(ggplot2)
library(haven)
library(cmdstanr)
library(localhierarchy)
```

## Load and Process Data

```{r load-data}
# Read data
data_folder <- "data_raw"
dat0 <- read_dta(here::here(data_folder, "ICEH_national.dta"))
regions_dat <- readr::read_csv(here::here(data_folder, "regions.csv"))

# Choose indicator
indicator_select <- "anc4"

# Process data
dat <- process_data(
  dat = dat0,
  regions_dat = regions_dat,
  indicator = indicator_select
)
```

## Set Validation Cutoff

Choose a cutoff year for validation. Data with `start_date >= validation_cutoff_year`
will be held out for validation.

```{r validation-cutoff}
# Set validation cutoff year
validation_cutoff_year <- 2018

# Check percentage of data held out
prop_held_out <- mean(dat$start_date >= validation_cutoff_year)
cat("Proportion of data held out:", round(prop_held_out, 3), "\n")
```

## Fit Model with Validation

```{r fit-validation}
fit_val <- fit_model(
  runstep = "step1ab",
  survey_df = dat,
  y = "invprobit_indicator",
  se = "se_invprobit_indicator",
  get_posteriors = TRUE,
  chains = 4,
  model_name = "rw2",
  validation_cutoff_year = validation_cutoff_year
)
```

## Visualize Validation Results

Plot the estimates to see how well the model predicts held-out data:

```{r plot-validation}
plots_val <- plot_estimates_local_all(
  results = fit_val,
  save_plots = TRUE,
  indicator_name = indicator_select
)
```

The plots will show:
- Fitted estimates using training data (before cutoff)
- Held-out observations (after cutoff) for visual comparison

## Comparing Multiple Cutoff Years

To assess robustness, run validation with different cutoff years:

```{r multiple-cutoffs}
cutoff_years <- c(2015, 2018, 2020)

for (cutoff in cutoff_years) {
  cat("\nFitting with cutoff year:", cutoff, "\n")
  cat("Data held out:", round(mean(dat$start_date >= cutoff), 3), "\n")

  fit <- fit_model(
    runstep = "step1ab",
    survey_df = dat,
    get_posteriors = TRUE,
    chains = 4,
    model_name = "rw2",
    runnumber = paste0("val_", cutoff),
    validation_cutoff_year = cutoff
  )
}
```

## Next Steps

- [Global Model Fitting](global_model_fitting.html): Full model fitting workflow
- [Model Diagnostics](diagnostics.html): Check convergence and model fit
